# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r, echo=TRUE}
act <- read.csv("activity.csv")
```



## What is mean total number of steps taken per day?
```{r, echo=TRUE}
library("ggplot2")
library("reshape2")
act.melt <- melt(act, id=c("date"), measure.vars=c("steps"),na.rm=TRUE)
steps.aggr <- dcast(act.melt, date ~ variable, sum)
qplot(steps, data=steps.aggr, binwidth = 500)

```



```{r mean.calcul, echo=TRUE}
mean.steps <- mean(steps.aggr$steps)
median.steps <- median(steps.aggr$steps)

```
The mean of the total number of steps taken each day is `r sprintf("%.2f",mean.steps)` and the mediam is `r sprintf("%.2f",median.steps)`.


## What is the average daily activity pattern?
```{r}
act.melt.inter <- melt(act, id=c("interval"), measure.vars=c("steps"), na.rm=TRUE)
steps.inter.aggr <- dcast(act.melt.inter, interval ~ variable, mean)
qplot(interval, steps, data = steps.inter.aggr, geom="line")
```


```{r}
max.inter <- steps.inter.aggr[which.max(steps.inter.aggr$steps),"interval"]
```
The interval `r sprintf("%.2f",max.inter)` contains the maximum number of steps


## Imputing missing values



```{r}
mis.val <- sum(is.na(act$steps))
```
1.The total number of rows with missing values is `r mis.val`


2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
Strategy: To use the mean for the 5-minute interval.

```{r}
act.melt.inter <- melt(act, id=c("interval"), measure.vars=c("steps"), na.rm=TRUE)
steps.inter.aggr <- dcast(act.melt.inter, interval ~ variable, mean)

head(steps.inter.aggr)
```




3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
Implementing strategy: Loping through the na values in the original dataset and finding for each interval the correpsonfing mean value in the dataset calculated in the previous point

```{r}
act.imp <- act
act.imp.na <- which(is.na(act.imp$steps))

for (i in act.imp.na) {
  int.na <- act.imp[i, "interval" ]
  tmp <- steps.inter.aggr[steps.inter.aggr$interval == int.na,]
  act.imp[i,1] <- tmp[1,2]
}

head(act.imp)
```


4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
library("ggplot2")
library("reshape2")
act.melt.imp <- melt(act.imp, id=c("date"), measure.vars=c("steps"),na.rm=TRUE)
steps.aggr.imp <- dcast(act.melt.imp, date ~ variable, sum)
qplot(steps, data=steps.aggr.imp, binwidth = 500)
```




```{r mean.calcul2, echo=FALSE}
mean.steps.imp <- mean(steps.aggr.imp$steps)
median.steps.imp <- median(steps.aggr.imp$steps)
```

The mean of the total number of steps taken each day is `r sprintf("%.2f",mean.steps.imp)`  and the mediam is `r sprintf("%.2f",median.steps.imp)`




## Are there differences in activity patterns between weekdays and weekends?
Yes, there are some differences in activity. During weekdays, there is just one peak, while on weekends there are several peaks. 

```{r}
act.imp$date <- as.Date(act.imp$date, "%Y-%m-%d")

act.imp$weekd <- sapply(weekdays(act.imp$date),switch,
  Monday="weekday",
  Tuesday="weekday",
  Wednesday="weekday",
  Thursday="weekday",
  Friday="weekday",
  Saturday="weekend",
  Sunday="weekend")


act.week <- melt(act.imp, id=c("interval","weekd"), measure.vars=c("steps"), na.rm=TRUE)
steps.week.aggr <- dcast(act.week, interval+weekd ~ variable, mean)

qplot(interval, steps, data = steps.week.aggr, geom="line")  + facet_wrap(~ weekd, nrow=2, ncol=1)
```
